Improving the Replicability of Psychological Science Through Pedagogy

### 2017-05-23 paper/paper.Rmd
paper for AMPPS submission

author-note:
    "*These authors contributed equally and are listed alphabetically. Thanks to the Stanford Department of Psychology and the Vice Provost for Graduate Education for funding to support the class. We are grateful to the authors of the original studies who provided materials and gave comments on a draft of this manuscript."

@@

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(xtable)
library(readr)
library(tidyr)
library(ggplot2)
library(BayesFactor)
library(ggrepel)
#devtools::install_github("langcog/langcog")
library(langcog)
library(cowplot)
library(png)
```

Replicability is a core value for empirical research and there is increasing concern throughout psychology that more independent replication is necessary [@osc2015;@wagenmakers2012]. Yet under the current incentive structure for science, replication is not typically valued. One potential solution to this problem is to make replication an explicit part of pedagogy: that is, to teach students about experimental methods by asking them to run replication studies [@frank2012;@grahe2012]. Despite enthusiasm for this idea [@lebel2015;@everett2015;@standing2016;@king2016], there is limited data beyond anecdotal reports or individual projects [@phillips2015;@lakens2013] to support its efficacy. And another sentence.

@@

We begin by describing the general method of the class, then discuss shared methods for the replications and the process of development for each project.

@@

While every course of this type will likely be tailored to the interests and instructional needs of a particular department and institution, there will also likely be some commonalities. To encourage others to share our materials, our course outline, project templates, and assignments are available publicly at [https://osf.io/98ta4](https://osf.io/98ta4). In addition, all of the most recent lecture slides and materials are avaialable on our course website at [http://psych254.stanford.edu](). We hope that others will share and reuse our materials as they consider how best to design a course customized to their context.

One of the barriers to more reproducible work in psychology is technical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web-based empirical studies or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

We end with a few recommendations for instructors designing such a course. First, plan as much as possible to make the project the centerpiece of the course, so that students begin the process as quickly as possible and do the maximal amount of skill learning in the context of their own goal-directed research. Second, rather than adopting the typical model of students turning in work at the end of the course and being assessed, consider a more collaborative model where the instructors provide iterative feedback on subcomponents of projects (e.g., sample analyses, parts of the experimental paradigm or materials). Such feedback can both save students time and frustration and provide a check on project quality before data are collected. Third and finally, try to share the excitement of discovery that is inherent in doing new research! We have found that the most rewarding part of teaching this class is the fun of getting to see the results of all the hard work at the end.

@@

In addition, our statistical findings mask a more optimistic message: In most of our projects, the next step for a motivated experimenter is clear. For example, in some projects we suspect that a followup could find strong evidence for the phenomenon of interest by titrating the particular planned analysis or the difficulty of the stimulus materials. In others, the difficulty appeared to be statistical power, since differences were in the predicted direction and sometimes reached significance in subsidiary analyses, so a followup would likely require a larger sample. And in some, differences of population would mean that followups would require further stimulus or task adaptation. More generally, we believe that our work here underscores the importance of iterated replication [@lewis2016].

>>>>

author-note:
    "*These authors contributed equally and are listed alphabetically. Thanks to the Stanford Department of Psychology and the Vice Provost for Graduate Education for funding to support the class. We are grateful to the authors of the original studies who provided materials and gave comments on a draft of this manuscript. RXDH, ENS, and MCF designed the project, supported data planning and data collection for all projects, analyzed the data, and wrote the paper. CA, JMA, RC, EH, MK, AL, SR, JR, SS, JS, and JT planned their individual projects, programmed their studies, collected their data, and analyzed their data."

@@

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
#devtools::install_github("kemacdonald/kmr")
library(readr)
library(dplyr)
library(xtable)
library(readr)
library(tidyr)
library(ggplot2)
library(BayesFactor)
library(ggrepel)
#devtools::install_github("langcog/langcog")
library(langcog)
library(cowplot)
library(png)
```

Replicability is a core value for empirical research and there is increasing concern throughout psychology that more independent replication is necessary [@osc2015;@wagenmakers2012]. Yet under the current incentive structure for science, replication is not typically valued. One potential solution to this problem is to make replication an explicit part of pedagogy: that is, to teach students about experimental methods by asking them to run replication studies [@frank2012;@grahe2012]. Despite enthusiasm for this idea [@lebel2015;@everett2015;@standing2016;@king2016], there is limited data beyond anecdotal reports or individual projects [@phillips2015;@lakens2013] to support its efficacy.

@@

We begin by describing the general method of the class, then discuss shared methods for the replications and the process of development for each project. In general, we report how we determined our sample size, all data exclusions, all manipulations, and all measures in the study.

@@

Though others' courses that incorporate replication projects will look quite different from our own, we recommend having the replication as a centerpiece of the course and giving ample opportunity for feedback throughout the study design, as these components were crucial to our success. To encourage others to share our materials, our course outline, project templates, and assignments are available publicly at [https://osf.io/98ta4](https://osf.io/98ta4). In addition, all of the most recent lecture slides and materials are avaialable on our course website at [http://psych254.stanford.edu](). We hope that others will share and reuse our materials as they consider how best to design a course customized to their context.

Finally, one of the barriers to more reproducible work in psychology is technical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web-based empirical studies or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

@@

In addition, our statistical findings mask a more optimistic message: In most of our projects, the next step for a motivated experimenter is clear. For example, in some projects we suspect that a followup could find strong evidence for the phenomenon of interest by titrating the particular planned analysis or the difficulty of the stimulus materials. In others, the difficulty appeared to be statistical power, since differences were in the predicted direction and sometimes reached significance in subsidiary analyses, so a followup would likely require a larger sample. And in some, differences of population would mean that followups would require further stimulus or task adaptation. More generally, we believe that our work here underscores the importance of iterated replication for pinpointing empirical effects and refining theories [see e.g., @lewis2016 for an example of this strategy and some discussion].
